---
title: "PSY304 stats demo"
author: "Shannon Dailey"
date: "10/29/2019"
output: pdf_document
---

```{r setup, echo=FALSE, include=FALSE}
library(knitr)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 5, fig.height = 4, fig.align='center')
data <- read_csv("demo_data.csv") %>% 
  mutate(gender = factor(gender),
         predom = factor(predom,
                            levels = c("cooperative", "associative", "parallel", "onlooker", "solitary", "unoccupied")))
```

# Descriptive Statistics
The dataset includes `r nrow(data)` participants total from 8 different cohorts (that is, 8 semesters of PSY304 students).

## Age x gender breakdown of participants
```{r}
ages_gender_table <- addmargins(table(data$age, data$gender), FUN = list(Total = sum), quiet = TRUE)
kable(ages_gender_table, col.names = c("Girls", "Boys","Total"), caption = "Age/gender breakdown of participants")
```

## Summary of dependent variables
```{r summary-tab, results='asis'}
library(fBasics)
tmp <- basicStats(data %>% dplyr::select(age, unoccupied, solitary, onlooker, parallel, associative, cooperative, soc_particip, bossy, pos_affect, awkward))
summary_table <- tmp[c("Mean", "Stdev", "Median", "Minimum", "Maximum"),] %>%
  t() %>%
  as.data.frame() %>% 
  mutate(Mean = paste(round(Mean,2), " (", round(Stdev,2), ")", sep = ""),) %>% 
  dplyr::select(-Stdev)
rownames(summary_table) <- c("Age", "Unoccupied", "Solitary play", "Onlooker", "Parallel play", "Associative play", "Cooperative play", "Social participation", "Bossiness", "Positive affect", "Social awkwardness")
kable(summary_table, caption = "Summary statistics of dependent variables")
```

\newpage

# Predominant play styles
A child's **predominant play style** is the type of play that child engaged in *the most* during the time sample.
```{r}
kable(table(data$predom), col.names = c("Type of play", "Count"), caption = "Number of children with each predominant play style")
```

## Predominant play style by age
```{r predom-age, fig.height=3.5, fig.width=4}
kable(table(data$predom, data$age), caption = "Predominant play style by age")

age_predom_table <- table(data %>% dplyr::select(age, predom))
library(broom)
chisq_age <- chisq.test(age_predom_table) %>% tidy()

ggplot(data %>% drop_na(predom), aes(x = age, fill = predom))+
  geom_bar(position = "fill")+
  labs(y = "Proportion of children", x = "Age (years)", fill = "Predominant play style", title = "Predominant play style by age")
```

**Chi-square test**: $\chi^2$(10) = `r round(chisq_age$statistic,2)`, *p* = `r round(chisq_age$p.value,2)`


## Predominant play style by gender
```{r}
kable(table(data$predom, data$gender), caption = "Predominant play style by gender")

ggplot(data %>% drop_na(predom), aes(x = gender, fill = predom))+
  geom_bar(position = "fill")+
  labs(y = "Proportion of children", x = "Gender", fill = "Predominant play style", title = "Predominant play by gender")

gender_predom_table <- table(data %>% dplyr::select(gender, predom))
chisq_gend <- chisq.test(gender_predom_table) %>% tidy()
```

**Chi-square test**: $\chi^2$(5) = `r round(chisq_gend$statistic,2)`, *p* = `r round(chisq_gend$p.value,2)`

\newpage
# Types of play
## Types of play by age
```{r}
props_long <- data %>%
  dplyr::select(child,age,gender,prop_unocc:prop_coop) %>% 
  pivot_longer(cols = prop_unocc:prop_coop, names_to = "type_play",
               names_prefix = "prop_", values_to = "prop") %>% 
  mutate(type_play = factor(type_play,
                            levels = c("coop", "assoc", "parallel", "onlooker", "solitary", "unocc")))

ggplot(props_long, aes(x = age, y = prop, color = type_play))+
  stat_summary(geom = "pointrange")+
  geom_smooth(method = "lm")+
  theme_bw()+
  labs(y = "Proportion of play", x = "Age  (years)", color = "Type of play", title = "Proportion of play types by  age")

aov_age_coop <- aov(prop ~ age, data = (props_long %>% dplyr::filter(type_play == "coop")))
#summary(aov_age_coop)

aov_age_solitary <- aov(prop ~ age, data = (props_long %>% dplyr::filter(type_play == "solitary")))
#summary(aov_age_solitary)
```

```{r}
age_by_type_tab_tmp1 <- props_long %>%
  drop_na(age) %>%
  group_by(type_play, age) %>%
  summarise(mean = mean(prop), sd = sd(prop)) %>%
  pivot_wider(names_from = age, values_from = c(mean, sd))

age_by_type_tab_tmp2 <- age_by_type_tab_tmp1

age_by_type_tab_tmp2$type_play <- c("Cooperative play", "Associative play", "Parallel play", "Onlooker", "Solitary play", "Unoccupied")

age_by_type_tab <- age_by_type_tab_tmp2 %>%
  dplyr::select(type_play, mean_3, sd_3, mean_4, sd_4, mean_5, sd_5)

library(kableExtra)
kable(age_by_type_tab,
      digits = 3,
      col.names = c("",
                "Mean","SD",
                "Mean", "SD",
                "Mean", "SD"),
      caption = "Proportion of play types by child age") %>% 
  kable_styling(bootstrap_options = "striped", latex_options = "hold_position") %>% 
  add_header_above(c(" " = 1, "3-year-olds" = 2, "4-year-olds" = 2, "5-year-olds" = 2))

library(papaja)
```

**ANOVA: Proportion of cooperative play by age**: `r apa_print(aov_age_coop)$statistic`

**ANOVA: Proportion of solitary play by age**: `r apa_print(aov_age_solitary)$statistic`

\newpage
## Types of play by gender
```{r type-gender, fig.width=4, fig.height=3}
props_gender <- props_long %>% group_by(gender,type_play) %>% summarise(mean_prop = mean(prop, na.rm = TRUE))

ggplot(props_long, aes(x = type_play, y = prop, color = gender))+
  stat_summary(geom = "pointrange", fun.data = "mean_cl_boot")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  labs(y = "Proportion of play", x = "Gender",
       color = "Type of play", title = "Types of play by gender")

t_gender_coop <- t.test(data = props_long %>% dplyr::filter(type_play == "coop"), prop ~ gender)
t_gender_solitary <- t.test(data = props_long %>% dplyr::filter(type_play == "solitary"), prop ~ gender)

type_by_gender_tab <- props_long %>%
  group_by(type_play, gender) %>%
  summarise(mean = mean(prop), sd = sd(prop)) %>%
  pivot_wider(names_from = gender, values_from = c(mean, sd)) %>% 
  dplyr::select(type_play, mean_female, sd_female, mean_male, sd_male)

type_by_gender_tab$type_play <- c("Cooperative play", "Associative play", "Parallel play", "Onlooker", "Solitary play", "Unoccupied")

kable(type_by_gender_tab,
      digits = 3,
      col.names = c("",
                "Mean","SD",
                "Mean", "SD"),
      caption = "Proportion of play types by child gender") %>% 
  kable_styling(bootstrap_options = "striped", latex_options = "hold_position") %>% 
  add_header_above(c(" " = 1, "Girls" = 2, "Boys" = 2))

```

**T-test: Cooperative play by gender**: `r apa_print(t_gender_coop)$statistic`

**T-test: Solitary play by gender**: `r apa_print(t_gender_solitary)$statistic`
